\section{Basic Obstacle Avoidance Problem}\label{s:BasicProblemDefinition}
    \noindent Given:

    \begin{enumerate}
        \item Initial system state $state_0$, for UAS model (eq. \ref{eq:vehicleModelAbstract}).
        
        \item Mission (eq. \ref{eq:missionAbstractSet}) to be executed.
        
        \item Space (eq. \ref{eq:SpaceDefinition}), with existing objects (eq. \ref{eq:ObjectDefinition}).
        
        \item Sensor system $\{sensor_1, sensor_2,\dots,sensor_k\}$ with \emph{sensor fusion} function (eq. \ref{eq:SensorFusionFunction}).
        
        \item Weather information (eq. \ref{eq:weatherProjection}) during the flight is available.
        
        \item Information sources $informationSource_1$, $informationSource_2$, $\dots$, $informationSource_l$ containing hard (eq. \ref{eq:hardSpaceConstraints}) and soft space constraints (eq. \ref{eq:softSpaceConstraints}),with existing data fusion function (eq. \ref{eq:DataFusionFunction}).
        
        \item \emph{Hard} (eq. \ref{eq:HFlightConstraints}) and \emph{Soft} (eq. \ref{eq:SFlightConstraints}) flight constraints given by ATM and rules of the air. 
    \end{enumerate}

    \noindent Generate \emph{Control command chain to} complete a mission (eq. \ref{eq:missionAbstractSet}) with the satisfaction of following conditions:
    
    \begin{enumerate}
        \item Waypoint passing function condition (eq. \ref{eq:waypointPassingFunction}).
        
        \item Set safety margin $s_m$ to $Occupied$  space in $KnownWorld(time)$ (eq. \ref{eq:knownWorldSet}) is not breached at any time (eq. \ref{eq:safetyMarginAbstract}).
    \end{enumerate}
