\cleardoublepage
\section{\secState{R}UAS Traffic Management}\label{sec:UASTrafficManagement}

\noindent The \emph{Traffic Management} for UAS is based on existing Air Traffic Management System for manned aviation \cite{icao4444}. The controlled airspace segments are \emph{static} and have one \emph{authority for one zone} principle. The dynamic zones have been proposed in \cite{gerdes2016dynamic}. However, it will be omitted for \emph{simplification purpose}. The necessity for \emph{UAS integration} into \emph{National Airspace} has been outlined in \cite{spriesterbach2013unmanned}.

The latest \emph{Airbus blueprint} \cite{airbusUTM2018blueprint} outlines some functionality. The main purpose of this section is to show \emph{Reach Set based Approach} capability to follow \emph{Usual Air Traffic Management} commands.

The \emph{section} is organized to introduce:
\begin{enumerate}
    \item \emph{UTM Architecture} (sec. \ref{sec:utmArchitecture}) - centralized ATM-like authority over airspace cluster.
    
    \item \emph{Handling Standard Collision Situations} - head-on approach (sec. \ref{sec:handlingHeadOnApproach}), converging situation (sec. \ref{sec:handlingConvergingManuever}), overtake (sec. \ref{sec:handlingOvertakeManuever}).
    
    \item \emph{Position Notification} (sec. \ref{sec:positionNotification}) - position notification design.
    
    \item \emph{Collision Case} (sec. \ref{sec:collisionCase}) - calculation and handling of \emph{collision situations}.
    
\end{enumerate}

\noindent The additional material can be found in:
\begin{enumerate}
	\item \emph{Cooperative Conflict Resolution} (app. \ref{sec:cooperativeConflictResolution}) - the model used for conflict resolution in \emph{controlled airspace}.
    
    \item \emph{Non-Cooperative Conflict Resolution} (app. \ref{sec:nonCooperativeConflictResolution})  - the model used for conflict resolution in \emph{non-controlled} airspace and \emph{emergency avoidance}.
    
    \item \emph{Weather Case} (app. \ref{sec:weatherCase}) - definition and handling of \emph{weather hazards}.
\end{enumerate}