\section{Linear Intersection}\label{s:linearIntersectionModel}
\paragraph{Idea:} There are \emph{small intruders} which have body \emph{smaller} than average $cell_{i,j,k}$ cell size. Its trajectory will stick to \emph{linear trajectory} prediction with high probability. 

\paragraph{Space Intersection Rate:} The \emph{Space Intersection Rate} for $cell_{i,j,k}$ is implemented as simple point cloud intersection. Where \emph{sufficiently thick} point cloud is defined along \emph{line} (eq. \ref{eq:linearModelIntruderLineIntersection}):

\begin{equation}\label{eq:linearModelIntruderLineIntersection}
    position(time) = position(time_0) + velocity \times time,\quad time \in [0,\infty[
\end{equation}

Then there exist projection function from local Euclidean  coordinates to local polar coordinates (eq. \ref{eq:projecionFunctionplanarEuclidlinearIntersection}. The function projects intruder trajectory (eq. \ref{eq:linearModelIntruderLineIntersection}) to planar coordinates $[distance,$ $horizontal^\circ,$ $vertical^\circ]$  as a set of sufficiently thick point cloud.

\begin{equation}\label{eq:projecionFunctionplanarEuclidlinearIntersection}
    polarSet:position(t)\to\{[distance, horizontal^\circ], vertical^\circ\}
\end{equation}


\noindent The \emph{space intersection rating} $SpaceIntersection(\circ)$ for line type is given as (eq. \ref{eq:baseIntersectionProbabilityLineIntersectionType}). If there exist non empty intersection of $polarSet\cap cell_{i,j,k}$ there is space intersection rate equal to  1, if intersection $polarSet\cap cell_{i,j,k} = \varnothing$ then the rate is zero.

\begin{equation}\label{eq:baseIntersectionProbabilityLineIntersectionType}
    space\left(\begin{gathered}Intruder,\\cell_{i,j,k}\end{gathered}\right)=
    \begin{cases}
        1:&\exists point \in polarSet(eq.\ref{eq:projecionFunctionplanarEuclidlinearIntersection}): point \in c_{i,j,k}\\
        0:&\text{otherwise}
    \end{cases}
\end{equation}

\begin{note}
    The \emph{intruder intersection rate} is multiplication of \emph{space intersection rate} and time intersection rate. The \emph{intersection rate} is calculated for \emph{every intruder} and \emph{selected intersection model} separately.
\end{note}
