\subsection{Data fusion}\label{s:sensorFusion}

\paragraph{Summary:} There is a need for the final threat assessment in the Avoidance Grid. The data fusion provides mechanisms to represent, process, and assess threat in the cell including the safety of trajectories in the RSA. The output of the data fusion procedure is used further in Avoidance run (sec. \ref{s:aviudabceGridRun}).


\paragraph{Introduction:} The data fusion interfaces \emph{Sensor Field} and \emph{Information Sources} from \emph{cell/trajectory properties}. The \emph{Data Fusion Function} is outlined in (\ref{eq:DataFusionFunction}). 

First, there will be an outline of \emph{Partial Rating} commutation. Then these ratings will be discredited into Boolean values as properties of \emph{Avoidance Grid/Trajectory}. Then these Boolean values will be used for further classification of  space into \emph{Free(t), Occupied(t), Restricted(t)} and \emph{Uncertain(t)}.

All mentioned ratings are the result of \emph{Filtered Sensor Readings} from \emph{Sensor Field} and \emph{Information Sources} with prior processing. This section will focus on \emph{final fuzzy value calculation} and \emph{discretization}. 
\begin{note}
    All rating values are in the \emph{range:} $[0,1]$, and they were introduced in previous sections.
\end{note}


\paragraph{Visibility:} The \emph{sensor reading} of \emph{sensor} if \emph{Sensor field} returns a value of \emph{visibility} for cell space in time of decision $t_i$.

The \emph{visibility} for the cell is given in (eq. \ref{eq:visibilityForCell}) as minimal visibility calculated from all capable sensors in \emph{Sensor Field}.

\begin{equation}\label{eq:visibilityForCell}
    visibility(cell_{i,j,k}) = \min \left\{\begin{aligned}visibility(cell_{i,j,k},&sensor_i):\\&\forall sensor_i \in Sensor Field\end{aligned}\right\}
\end{equation}

\noindent The example of \emph{visibility} calculation for \emph{LiDAR} sensor is given in (fig. \ref{fig:mapObstacleStatesAfterDataFusion}).

\begin{note}
    Sensor reliability for \emph{visibility} is already accounted for prior \emph{data fusion}. If not \emph{weighted average} should be used instead. 
\end{note}

\paragraph{Detected Obstacle:} Sensors detect the physical obstacles  in \emph{Sensor Field}. Each \emph{sensor} returns \emph{detected obstacle rating} in the range $[0,1]$ reflecting the probability of obstacle occurrence in a given  cell.

The \emph{maximal value} of \emph{detected obstacle} rating is selected from readings multiplied by \emph{visibility rating} to enforce \emph{visibility bias}.

\begin{multline}\label{eq:detectedObstacleRatingForCell}
    obstacle(cell_{i,j,k}) = \max \left\{\begin{aligned}obstacle(cell_{i,j,k},&sensor_i):\\&\forall sensor_i \in SensorField\end{aligned}\right\}\times\dots\\\dots\times visibility(cell_{i,j,k})
\end{multline}

\noindent The example of \emph{detected obstacle rating} calculation for \emph{LiDAR} sensor is given in (eq. \ref{eq:naiveObstacleRate}).

\paragraph{Map Obstacle:} The \emph{Information Sources} are feeding \emph{Avoidance Grid} with partial information of \emph{Map obstacle rating}. \emph{Map Obstacle Rating} shows the certainty that \emph{charted obstacle} is in a given cell. This property is bound to \emph{Information Source}, and it has the \emph{range} in  $[0,1]$.

The \emph{Map Obstacle Rating} for a cell (eq. \ref{eq:mapObstacleRatingForCell}) is calculated as the product of maximal \emph{Map Obstacle Rating} and \emph{inverse visibility}. This gives \emph{visibility biased} certainty of \emph{Map Obstacle}.

\begin{multline}\label{eq:mapObstacleRatingForCell}
    map(cell_{i,j,k}) = \max 
    \left\{\begin{aligned}map(&cell_{i,j,k},source_i):\\&\forall source_i \in InformationSources\end{aligned}\right\}\times\dots\\\dots\times \left(1-visibility(cell_{i,j,k})\right)
\end{multline}

\noindent The example of \emph{Map Obstacle Rating} calculation is given in (fig. \ref{fig:mapObstacleStatesAfterDataFusion}).


\paragraph{Intruder:} There is a set of \emph{Active Intruders}, each intruder is using its \emph{parametric intersection model}. This parametric \emph{intersection} model calculates \emph{partial intersection ratings} representing \emph{intersection certainty} ranging in $[0,1]$. The more \emph{partial intersection rating} is closer to 1 the higher is the probability of aerial collision with that intruder in that cell. 

The \emph{geometrical bias} is used for cumulative of multiple intruders; the \emph{intruders are not cooperative}; therefore their occurrence cannot be addressed by the simple \emph{maximum}. The proposed formula (eq. \ref{eq:intruderRatingForCell}) is simply bypassing the intruder rating if there is one intruder. If there  are more intruders, the geometrical bias is applied.


\begin{equation}\label{eq:intruderRatingForCell}
    intruder(cell_{i,j,k}) = 1 - \prod_{\forall intruder_i \in Intruders} \left(1- intersection\left(\begin{gathered}cell_{i,j,k},\\intruder_i\end{gathered}\right)\right)
\end{equation}

\noindent The \emph{intruder intersection models} are outlined in (app. \ref{app:IntruderProbabilisticModels}). 

\paragraph{Constraint:} The \emph{constraints} are coming from various \emph{Information Sources}, the \emph{hierarchical constraint application} is resolved by higher level logic. All \emph{constraints} in this context are considered as \emph{hard}.

The \emph{Constraints rating} (eq. \ref{eq:constraintRatingForCell}) is in the \emph{range} $[0,1]$ reflecting certainty of constraint application in the cell (usually 1).

\begin{equation}\label{eq:constraintRatingForCell}
    constraint(cell_{i,j,k}) = \max \left\{\begin{aligned}constraint(&cell_{i,j,k},source_i):\\&\forall source_i \in InformationSources\end{aligned}\right\}
\end{equation}

\noindent The \emph{Constraint Rating} calculation example for \emph{static} constraints is given in (sec. \ref{s:virtualConstraints}), the example for \emph{moving} constraints is given by (def. \ref{def:movingConstraint}).

\begin{note}{Weather}
    is already considered in constraints; the weather is handled as soft/hard static/moving constraints.
\end{note}

\paragraph{Threat:} The concept of threat is a \emph{rating of expected harm} to receive in a given portion of space. The threat can be time-bound to \emph{decision time $t_i$} (time sensitive \emph{intruder intersection models}).

The \emph{harm prioritization} is addressed by higher navigation logic (fig. \ref{fig:missionControlRunActivityDiagram}). All \emph{sources of harm} are considered as equal. The threat is formalized in the \emph{following definition}:

\begin{definition}{The Threat}\label{def:threat} is considered as any source of harm. The threat is a \emph{maximal aggregation} of various harm ratings. Our \emph{threat} for a  specific cell is defined by (eq. \ref{eq:threatRatingForCell}).
    \begin{equation}\label{eq:threatRatingForCell}
        threat(cell_{i,j,k}) = \max\left\{\begin{gathered}obstacle(cell_{i,j,k}),map(cell_{i,j,k}),\\intruder(cell_{i,j,k}),constraint(cell_{i,j,k})\end{gathered}\right\}
    \end{equation}
\end{definition}

\paragraph{Reachability:} The \emph{Reachability} for trajectory reflects how safe is the \emph{path along}. The \emph{Threat} (def. \ref{def:threat}) for each cell has been already assessed.  The set of \emph{Passing Cells} is defined in \emph{Trajectory Footprint} (eq. \ref{eq:setOfPassedCells}).

The \emph{Trajectory Reachability} is given as a product of \emph{Threats} along the trajectory (eq. \ref{eq:trajectoryReachibility}). The \emph{Trajectory Reachability} can be calculated for each \emph{trajectory segment} given as $\{movement_1,\dots,movement_i\}$ $\subset$ $Buffer$ originating from $state_0$.


\begin{equation}\label{eq:trajectoryReachibility}
    reachibility(Trajectory) = \prod_{Passing Cells}^{\forall cell_{i,j,k}\in} \left(1- threat(c_{i,j,k})\right)
\end{equation}

\begin{note}
    The \emph{Reachability} of \emph{trajectory} segment gives the property of \emph{safety} of route from the beginning, until the last point of the segment. There can be a very unsafe trajectory which is very safe from the beginning.
\end{note}


The \emph{Reachability} of the \emph{cell} is given by the best trajectory segment passing through the \emph{given cell}. This is given by property, that every trajectory is originating from root $state_0$, which means that one safe route is sufficient to reach space in the cell.

The \emph{Trajectory segment} reachability is sufficient, because the overall performance is not interesting, the \emph{local reachability} is sufficient. The cell reachibility is formally defined in (eq. \ref{eq:cellReachibility}).

\begin{multline}\label{eq:cellReachibility}
    reachability(cell_{i,j,k}) = \max\{Trajectory.Segment(cell_{i,j,k}). Reachability: \\\forall Trajectory \in Passing Trajectories (cell{i,j,k})\}
\end{multline}
    
\begin{note}
    Function Trajectory.Segment($cell_{i,j,k}$). Reachability gives same results for any segment in $cell_{i,j,k}$, because (eq. \ref{eq:trajectoryReachibility}) accounts each cell $threat$ only once.
\end{note}

\paragraph{Discretization:} The \emph{fault tolerant} implementation needs to implement sharp Boolean values of properties mentioned before. The \emph{fuzzy values} are usually threshold to Boolean equivalent. The \emph{operational standards} for \emph{Manned Aviation} \cite{icao4444} demands the fail rate below $10^{-7}$ because there is no definition for \emph{UAS} the \emph{minimal fail rate} is expected to be at a similar level.

The \emph{fuzzy values} $[0,1]$ are projected to \emph{Boolean} properties of \emph{cell} and \emph{Trajectory} in the following manner (tab. \ref{tab:defuzificationRatings}).


The high values of \emph{Visibility} (eq. \ref{eq:visibilityForCell}) and \emph{Reachability} (eq. \ref{eq:cellReachibility}, \ref{eq:trajectoryReachibility}) are expected. The low \emph{threshold} for \emph{threats} values is expected. The error margin is solved by \emph{Sensor Fusion}, therefore, initial \emph{false positive} cases have a low rate. The \emph{Detected Obstacle Rate} (eq. \ref{eq:detectedObstacleRatingForCell}), \emph{Map Obstacle Rate} (eq. \ref{eq:mapObstacleRatingForCell}), \emph{Intruder Rate} (eq. \ref{eq:intruderRatingForCell}), and \emph{Constraint Rate} (eq. \ref{eq:constraintRatingForCell}) thresholds are considered low.

\begin{table}[H]
    \centering
    \begin{tabular}{c|ccc}
        \multicolumn{4}{c}{Threshold = $10^{-7}$}\\\hline\hline
        Visibile & $visibility(cell_{i,j,k})$&$\ge$&$(1-threshold)$ \\\hline
        Detected Obstacle &  $obstacle(cell_{i,j,k}) $&$ \ge $&$ threshold$\\\hline
        Map Obstacle &  $map(cell_{i,j,k})$&$\ge$&$threshold$\\\hline
        Intruder &  $intruder(cell_{i,j,k})$&$\ge$&$threshold$\\\hline
        Constraint &  $constraint(cell_{i,j,k})$&$\ge$&$threshold$\\\hline\hline
        Reachable Trajectory &  $reachability(trajectory)$&$\ge$&$(1-threshold)$\\\hline
        Reachable Cell &  $reachibility(cell_{i,j,k})$&$\ge$&$(1-threshold)$
    \end{tabular}
    \caption{Changing ratings from fuzzy to Boolean parameters.}
    \label{tab:defuzificationRatings}
\end{table}


\paragraph{Space Classification:} The \emph{Data Fusion Function} is outlined in (\ref{eq:DataFusionFunction}). This classification is resulting in four distinct cell sets.

The \emph{Uncertain} space for decision time $t_i$ is a portion of \emph{Avoidance Grid} which \emph{UAS} cannot \emph{read} with \emph{Sensor Field}. The \emph{cells} with a $\neg Visible$ property. The \emph{Uncertain} space is given by (eq. \ref{eq:UncertainDataFusion}).

\begin{equation}\label{eq:UncertainDataFusion}
    Uncertain(t_i) = \left\{cell_{i,j,k}:cell_{i,j,k}\in AvoidanceGrid(t_i),cell_{i,j,k}.\neg Visible \right\}
\end{equation}

\noindent The \emph{Occupied} space for decision time $t_i$ is the set of cells which are classified as \emph{Detected Obstacles}. The \emph{Visibility} is not an issue, due to the initial damping in (eq. \ref{eq:detectedObstacleRatingForCell}). The formal definition is the space portion where it is possible to detect \emph{obstacle bodies} or their portions (eq. \ref{eq:ocuupiedDataFusion}).

\begin{equation}\label{eq:ocuupiedDataFusion}
    Occupied(t_i) = \left\{cell_{i,j,k}:\begin{aligned}&cell_{i,j,k}\in AvoidanceGrid(t_i),\\&cell_{i,j,k}.DetectedObstacle\end{aligned}\right\}
\end{equation}

\noindent The \emph{Constrained} space for decision time $t_i$ is \emph{Visible} portion of \emph{Avoidance Grid} where the \emph{Intruder} or \emph{Constraint} is present. The mathematical formulation is given in (eq. \ref{eq:constrainedDataFusion}).

\begin{equation}\label{eq:constrainedDataFusion}
    Constrained(t_i) = \left\{cell_{i,j,k}:
    \begin{aligned}
        &cell_{i,j,k} \in AvoidanceGrid(t_i),\\
        &cell_{i,j,k}.Visible,\\
        &cell_{i,j,k}.Constraint \vee cell_{i,j,k}.Intruder
    \end{aligned}\right\}
\end{equation}

\noindent The \emph{Free} space is the space which is \emph{Visible} and $\neg Obstacle$,  $\neg Intruder$, and, $\neg Constrained$. The mathematical definition is simple set subtractions from \emph{Avoidance Grid} (eq. \ref{eq:freeDataFusion}).

\begin{multline}\label{eq:freeDataFusion}
    Free(t_i) = AvoidanceGrid(t_i) -\dots\\\dots -\left(Uncertain(t_i)\cup Occupied(t_i)\cup  Constrained(t_i)\right)
\end{multline}

\noindent The \emph{Reachable} space for time $t_i$, used in \emph{Avoidance} because its free and there is a safe trajectory, is given as a set of cells from \emph{Avoidance Grid} which are \emph{Reachable}. The mathematical definition is given in (eq. \ref{eq:ReachableDataFusion}).

\begin{equation}\label{eq:ReachableDataFusion}
    Reachable(t_i) = \left\{cell_{i,j,k}:\begin{aligned}&cell_{i,j,k}\in AvoidanceGrid(t_i),\\&cell_{i,j,k}.Reachable\end{aligned}\right\}
\end{equation}

\begin{note}{The Reachable Space at decision time $t_i$:} 
The \emph{Reachable space} is a non-empty set and its a subset of \emph{Free($t_i$)} space:    

\begin{equation}\label{eq:reachableDataFusionConstraints}
    |Reachable(t_i)| > 0, \quad Reachable(t_i) \subset Free(t)
\end{equation}
\end{note}
